---
title: "Assignment_3_Script"
author: "Hayden Simm"
date: "2026-01-20"
output: html_document
---

JD: It's hard to be sure what's new here. Maybe better to respond to Ben back in Assignment 2, and start this assignment by reading an .rds file. You can also use the README to tell us how things go together.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# loading packages --------------------------------------------------------
suppressPackageStartupMessages(
  library(tidyverse)
  )
suppressPackageStartupMessages(
  library(respR)
  )
```

creating the dataframe
```{r}
df_respirometry_acclimation <- read.delim("KILLIFISH_RESPIROMETRY_2_ACCLIMATION.txt")
df_respirometry_opposite <- read.delim("KILLIFISH_RESPIROMETRY_2_OPPOSITE.txt")
```
**BMB**: is there any way you can get these data in a saner format?
**response**: the software used to collect this info is called "pyroscience" 
  and the file outputs are supposed to be their own file types ".pyr", so yes 
  it is a mess and part of the reason I'm here!

Removing rows by a row name in the software instead of manually
```{r}
start_row_acclimation <- which(apply(df_respirometry_acclimation, 1, 
  function(x) any(x == "#--- Measurement Data --------------------------------------------------------------------"))) 
df_respirometry_acclimation <- df_respirometry_acclimation[(start_row_acclimation + 1):nrow(df_respirometry_acclimation), ]
colnames(df_respirometry_acclimation) <- df_respirometry_acclimation[1, ]
df_respirometry_acclimation <- df_respirometry_acclimation [-1,]

start_row_opposite <- which(apply(df_respirometry_opposite, 1, 
  function(x) any(x == "#--- Measurement Data --------------------------------------------------------------------"))) 
df_respirometry_opposite <- df_respirometry_opposite[(start_row_opposite + 1):nrow(df_respirometry_opposite), ]
colnames(df_respirometry_opposite) <- df_respirometry_opposite[1, ]
df_respirometry_opposite <- df_respirometry_opposite [-1,]
```

Setting up a function to process the dataframes for both acclimation and opposite
```{r}
process_respirometry_df <- function(df) {
  df %>%
    select(
    " dt (s) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.1 Main]",
    "Oxygen (%air sat.) [A Ch.2 Main]",
    "Oxygen (%air sat.) [A Ch.3 Main]",
    "Oxygen (%air sat.) [A Ch.4 Main]",
    "Oxygen (%air sat.) [B Ch.1 Main]",
    "Oxygen (%air sat.) [B Ch.2 Main]",
    "Oxygen (%air sat.) [B Ch.3 Main]",
    "Oxygen (%air sat.) [B Ch.4 Main]"
    ) %>%
    janitor::clean_names(case = "snake") %>%
    mutate(across(everything(), as.numeric)) %>%
    drop_na()
  }
```

Running the function
```{r}
df_respirometry_acclimation <- process_respirometry_df(df_respirometry_acclimation)
df_respirometry_opposite <- process_respirometry_df(df_respirometry_opposite)
```

**Note**: I recognize that have each of the following respR functions separate is redundant when I could define it as a function and run it repeatedly like above. However the individual variance between chambers makes this tricky because the "weight", start", "wait" and "measure" times all slightly differ so I decided to do this manually. 
**as well**: to respond to a comment from assignment 2, I agree that the automation of this calculation is tricky. The historical way of process data this was to literally take the slope by hand from the raw file. Our lab is experimenting with trying to use this Rpackage "respR" instead, and I have intentionally chosen this data to bring to class to get the opportunity to play with the data and see if it is something worth using. 
```{r}
A1_acclimation<-df_respirometry_acclimation %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_1_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 200, 
                measure = 300, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 75, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 28, t = 10, mass = 0.006843) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
A2_acclimation<-df_respirometry_acclimation %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_2_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 200, 
                measure = 300, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 100, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 28, t = 10, mass = 0.008887) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
A3_acclimation<-df_respirometry_acclimation %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_3_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 200, 
                measure = 300, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 100, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 10, mass = 0.006597) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
A4_acclimation<-df_respirometry_acclimation %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_a_ch_4_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 200, 
                measure = 300, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 150, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 10, mass = 0.007571) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
B1_opposite<-df_respirometry_opposite %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_1_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 100, 
                measure = 100, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 50, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 28, t = 24, mass = 0.006843) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
B2_opposite<-df_respirometry_opposite %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_2_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 100, 
                measure = 150, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 100, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 28, t = 24, mass = 0.008887) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
B3_opposite<-df_respirometry_opposite %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_3_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 100, 
                measure = 180, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 75, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 24, mass = 0.006597) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

```{r}
B4_opposite<-df_respirometry_opposite %>%
  select("dt_s_a_ch_1_main", "oxygen_percent_air_sat_b_ch_4_main") %>%
  na.omit() %>%
  inspect(time = 1, oxygen = 2, plot = FALSE) %>%
  subset_data(from = 30000, to = 40000, by = "time") %>% #subset to after MMR and middle of night 
  auto_rate.int(start = 600, #time that flush pump is on, in seconds
                wait = 100, 
                measure = 180, #wait + measure times should add to time that flush pump is off
                method = "linear", width = 75, by = "time") %>%
  convert_rate(oxy.unit = '%Air', time.unit = 's', output.unit = 'umol/min/g', 
               volume = 0.175, S = 11, t = 24, mass = 0.007571) %>%
  #plot(pos =10) %>%
  #summary(pos =10) %>% this is to specifically look at specific replicates
  select_rate(method = "rsq", n = c(0.9,1)) %>%
  select_rate(method = "lowest_percentile", n = 0.1) %>%
  mean() %>% #this is your final value
  summary(export = TRUE)
```

JD: Not sure what to make of all the plots above here. They could have suppressed, or the flow explained a bit better, or both.

listing dataframes to combine them
```{r}
df_list <- list(
  A1_acclimation,
  A2_acclimation,
  A3_acclimation,
  A4_acclimation,
  B1_opposite,
  B2_opposite,
  B3_opposite,
  B4_opposite
)

df_combined <- bind_rows(df_list)
```

JD: It's not clear why you don't just make t into a factor (or put the factor in your aes() statement). Also not clear why you use different syntax (mutate vs. $) to do almost exactly the same thing to t and S

Factoring to group by for plots
```{r}
df_combined <- df_combined %>%
  mutate(temp_group = factor(t))
df_combined$S <- factor(df_combined$S)
```


JD: Why are you taking absolute values? That's generally a strange move, and worth explaining here. Since they're all presumably supposed to be negative using "-" is probably the more principled approach (although I need to know more to be sure). It's also not clear why you need to do so much manual stuff, nor why it's clear to make a boxplot for so few points.

Boxplot
```{r}
ggplot(df_combined, aes(x = temp_group, y=abs(rate.output))) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(
    aes(shape = S, fill = S),
    width = 0.1,
    size = 3,
    color = "black"
  ) +
  scale_shape_manual(
    values = c("11" = 21, "28" = 24),
    name = "Salinity (ppt)"
  ) +
  scale_fill_manual(
    values = c("11" = "white", "28" = "black"),
    name = "Salinity (ppt)"
  ) +
  scale_y_log10() +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(color = "grey75"),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = "Assay Temperature (\u00B0C)",
    y =  expression("Metabolic Rate (" * mu * "mol O"[2] ~ min^{-1} ~ g^{-1} * ")")
  )
```

JD: I made this just to try to understand the data, but it seems to cover what we're seeing in the plot above
```{r scatter}
## summary(df_combined)
print(ggplot(df_combined)
	+ aes(x = temp_group, y=-rate.output, color=S, shape=S)
	+ geom_point(position=position_dodge(width=0.5), size=2)
	+ theme_classic() ## Consider using theme_set at top
)
```

JD: I can't tell at all what the Area plot is supposed to do. You say it's not a good method, but I'm wondering if it's even implemented in a sensible way.

Area plot
```{r}
df_bubbles <- df_combined %>%
  mutate(S = factor(S)) %>%
  group_by(S) %>%
  summarise(
    mean_rate = mean(abs(rate.output), na.rm = TRUE)
  )
ggplot(df_bubbles, aes(x = S, y = 1)) +
  geom_point(
    aes(size = mean_rate, fill = S),
    shape = 21,
    color = "black",
    stroke = 1.2
  ) +
  scale_size_area(max_size = 20) +
  scale_fill_manual(
    values = c(
      "11" = "grey90",
      "28" = "grey65"
    ),
    name = "Salinity (ppt)"
  ) +
  guides(fill = "none") +
  theme_classic() +
  labs(
    x = "Salinity (ppt)",
    y = NULL,
    size = "Metabolic Rate\n(µmol O₂·min⁻¹·g⁻¹)"
  ) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y  = element_blank()
  )
```

Time series plot
```{r}
df_acc_plot <- df_respirometry_acclimation %>%
  slice(150:400) %>%
  mutate(t_align = dt_s_a_ch_1_main - first(dt_s_a_ch_1_main))

df_opp_plot <- df_respirometry_opposite %>%
  slice(100:350) %>%
  mutate(t_align = dt_s_a_ch_1_main - first(dt_s_a_ch_1_main))

ggplot() +
  geom_line(
    data = df_acc_plot,
    aes(x = t_align, y = oxygen_percent_air_sat_a_ch_1_main),
    color = "blue",
    linewidth = 1.2
  ) +
  geom_line(
    data = df_opp_plot,
    aes(x = t_align, y = oxygen_percent_air_sat_a_ch_1_main),
    color = "red",
    linewidth = 1.2
  ) +
  geom_text(
    data = df_acc_plot %>% slice(n()),
    aes(x = t_align, y = oxygen_percent_air_sat_a_ch_1_main),
    label = "10\u00B0C Acclimation",
    color = "blue",
    size = 5,
    hjust = -0.1
  ) +
  geom_text(
    data = df_opp_plot %>% slice(n()),
    aes(x = t_align, y = oxygen_percent_air_sat_a_ch_1_main),
    label = "24\u00B0C Acclimation",
    color = "red",
    size = 5,
    hjust = -0.1,
    vjust = 0.05
  ) +
  coord_cartesian(clip = "off") +
  theme_classic() +
  theme(
    plot.margin = margin(t = 5, r = 110, b = 5, l = 5),
    panel.grid.major = element_line(color = "grey70"),
    panel.grid.minor = element_line(color = "grey85")
  ) +
  labs(
    x = expression(Delta*"Time (s)"),
    y = "Oxygen Percentage (%)"
  )
```
**descriptions of plots are in read me**

Grade: 1.8/3
